IML Term project

### 2022-12-11 main_markdown.rmd
trees

# 168
qda_acc.168.pca <- data.frame(CV = accuracy(qda.pred.cv.168.pca,
                                            npf.168.pca$class2),
                              LOOCV = accuracy(qda.pred.loocv.168.pca, 
                                               npf.168.pca$class2))
qda_perp.168.pca <- data.frame(CV = perplexity(qda.pred.cv.168.pca, 
                                               npf.168.pca$class2),
                               LOOCV = perplexity(qda.pred.loocv.168.pca, 
                                                  npf.168.pca$class2))
kable(cbind(qda_acc.168.pca, qda_perp.168.pca)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2)) %>%
  kable_styling(latex_options = "HOLD_position")
```
## Multiple classes

>>>>

# 168
qda_acc.168.pca <- data.frame(CV = accuracy(qda.pred.cv.168.pca,
                                            npf.168.pca$class2),
                              LOOCV = accuracy(qda.pred.loocv.168.pca, 
                                               npf.168.pca$class2))
qda_perp.168.pca <- data.frame(CV = perplexity(qda.pred.cv.168.pca, 
                                               npf.168.pca$class2),
                               LOOCV = perplexity(qda.pred.loocv.168.pca, 
                                                  npf.168.pca$class2))
kable(cbind(qda_acc.168.pca, qda_perp.168.pca)) %>%
  add_header_above(c("Accuracy" = 2, "Perplexity" = 2)) %>%
  kable_styling(latex_options = "HOLD_position")
```

### Tree-based methods
```{r tree, warning=FALSE, message=FALSE}
set.seed(68)
library(tree)
library(ISLR2)
idx1 <- sample(nrow(npf.fact), nrow(npf.fact)/2)
npf.fact.train <- npf.fact[idx1, ]
npf.fact.test <- npf.fact[-idx1, ]
idx2 <- sample(nrow(npf.168.fact), nrow(npf.168.fact)/2)
npf.168.fact.train <- npf.168.fact[idx2, ]
npf.168.fact.test <- npf.168.fact[-idx2, ]

tree.train.fact <- tree(class2 ~ ., npf.fact.train)
summary(tree.train.fact)
tree.train.168.fact <- tree(class2 ~ ., npf.168.fact.train)
summary(tree.train.168.fact)

par(mfrow = c(1, 2))
plot(tree.train.fact)
plot(tree.train.168.fact)

prune.train.fact <- prune.misclass(tree.train.fact, best = 8)
prune.train.168.fact <- prune.misclass(tree.train.168.fact, best = 8)
par(mfrow = c(1, 2))
plot(prune.train.fact)
plot(prune.train.168.fact)

tree.pred.fact <- predict(prune.train.fact, npf.fact.test)
tree.pred.168.fact <- predict(prune.train.168.fact, npf.168.fact.test)
LOOCV.tree = loocv(prune.train.fact, data = npf.fact.test)
LOOCV.168.tree = loocv(prune.train.168.fact, data = npf.168.fact.test)
CV.tree <- cv(prune.train.fact, data = npf.fact.test)
CV.168.tree = cv(prune.train.168.fact, data = npf.168.fact.test)

acc.tree <- data.frame(original = mean(tree.pred.fact == npf.fact.test$class2),
                           CV = mean(CV.tree == npf.fact.test$class2),
                           LOOCV = mean(LOOCV.tree == npf.fact.test$class2))
acc.tree.168 <- data.frame(original = mean(tree.pred.168.fact == npf.168.fact.test$class2),
                           CV = mean(CV.168.tree == npf.168.fact.test$class2),
                           LOOCV = mean(LOOCV.168.tree == npf.168.fact.test$class2))
per.tree <- data.frame(original = perplexity(tree.pred.fact, npf.fact.test$class2),
                           CV = perplexity(CV.tree, npf.fact.test$class2),
                           LOOCV = perplexity(LOOCV.tree, npf.fact.test$class2))
per.tree.168 <- data.frame(original = perplexity(tree.pred.168.fact, npf.168.fact.test$class2),
                           CV = perplexity(CV.168.tree, npf.168.fact.test$class2),
                           LOOCV = perplexity(LOOCV.168.tree, npf.168.fact.test$class2))

kable(cbind(rbind(acc.tree, acc.tree.168), rbind(per.tree, per.tree.168))) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3)) %>%
  kable_styling(latex_options = "HOLD_position")
```

Random forest tested with original and 168 dataset, and accuracies and perplexities reported.
Why is mtry 12?

```{r randomForest, warning=FALSE, message=FALSE}
library(randomForest)
set.seed(1)

# Select training and test data
idx <- sample(nrow(npf.168), nrow(npf.168)/2)
npf_train <- npf[idx, ]
npf_test <- npf[-idx, ]
npf_train168 <- npf.168[idx, ]
npf_test168 <- npf.168[-idx, ]

npf.RF.mod <- randomForest(class2 ~ ., data = npf_train, mtry = 12, importance = TRUE)
npf.RF.pred <- predict(npf.RF.mod, newdata = npf_test)
npf.RF.CV <- cv(npf.RF.mod, data = npf_test)
npf.RF.LOOCV <- loocv(npf.RF.mod, data = npf_test)

npf168.RF.mod <- randomForest(class2 ~ ., data = npf_train168, mtry = 12, importance = TRUE)
npf168.RF.pred <- predict(npf168.RF.mod, newdata = npf_test168)
npf168.RF.CV <- cv(npf168.RF.mod, data = npf_test168)
npf168.RF.LOOCV <- loocv(npf168.RF.mod, data = npf_test168)

yhat.RF <- predict(npf.RF.mod, newdata = npf_test)
plot(yhat.RF, npf_test$class2)

accuracies <- data.frame(original = accuracy(npf.RF.pred, npf_test$class2),
                           CV = accuracy(npf.RF.CV, npf_test$class2),
                           LOOCV = 0)
accuracies.168 <- data.frame(original = accuracy(npf168.RF.pred, npf_test168$class2),
                           CV = accuracy(npf168.RF.CV, npf_test168$class2),
                           LOOCV = accuracy(npf168.RF.LOOCV, npf_test168$class2))
perplexities <- data.frame(original = perplexity(npf.RF.pred, npf_test$class2),
                           CV = perplexity(npf.RF.CV, npf_test$class2),
                           LOOCV = 0)
perplexities.168 <- data.frame(original = perplexity(npf168.RF.pred, npf_test168$class2),
                           CV = perplexity(npf168.RF.CV, npf_test168$class2),
                           LOOCV = perplexity(npf168.RF.LOOCV, npf_test168$class2))
table <- cbind(rbind(accuracies, accuracies.168), rbind(perplexities, perplexities.168))

knitr::kable(table, digits = 3) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3)) %>%
  kable_styling(latex_options = "HOLD_position")

## Multiple classes
